A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium} symbioses and coral bleaching as an alternate stable state

### 2016-11-02 JTB_ms/JTB_ms.Rmd
update Eq. nos.

| Symbol  | Description                            | Units               | Eq. no. |
| ------- | -------------------------------------- | ------------------- | ------- |
| $j_X$     | Prey uptake rate                       | molX CmolH^-1^ d^-1^ | 3 |
| $j_N$     | Nitrogen uptake rate                   | molN CmolH^-1^ d^-1^ | 4 |
| $j_{HG}$  | Host biomass formation rate            | CmolH CmolH^-1^ d^-1^| 5 |
| $r_{NH}$  | Recycled nitrogen from host turnover   | molN CmolH^-1^ d^-1^ | 6 |
| $\rho_N$  | Nitrogen shared with the symbiont      | molN CmolH^-1^ d^-1^ | 7 |
| $j_{eC}$  | Excess carbon used to activate host CCMs| molC CmolH^-1^ d^-1^ |  |
| $j_L$     | Light absorption rate                  | molph CmolS^-1^ d^-1^| 9 |
| $j_{CO_2}$| CO~2~ input to photosynthesis          | molCO~2~ CmolH^-1^ d^-1^| 10 |
| $r_{CH}$  | Recycled CO~2~ from host turnover      | molCO~2~ CmolH^-1^ d^-1^| 11 |
| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | molCO~2~ CmolS^-1^ d^-1^| 12 |
| $j_{CP}$  | Photosynthesis rate                    | molC CmolS^-1^ d^-1^ | 13 |
| $j_{eL}$  | Light energy in excess of photochemistry| molph CmolS^-1^ d^-1^| 14 |
| $j_{NPQ}$ | Total capacity of NPQ                  | molph CmolS^-1^ d^-1^ | 15 |
| $c_{ROS}$ | ROS production proportional to baseline | -- | 16 |
| $r_{NS}$  | Recycled nitrogen from symbiont turnover| molN CmolS^-1^ d^-1^| 17 |
| $j_{SG}$  | Symbiont biomass formation rate        | CmolS CmolS^-1^ d^-1^| 18 |
| $\rho_C$  | Fixed carbon shared with host          | molC CmolS^-1^ d^-1^ | 19 |
| $j_{ST}$  | Symbiont turnover rate                 | CmolS CmolS^-1^ d^-1^| 20 |

Table 2. Model parameters

@@

where $\rho_C$ is fixed carbon shared by the symbiont (see Eq. XX), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover:

@@

CO~2~ arrives at the photosynthesis SU from multiple sources: in addition to the CO~2~ actively supplied by the host through its CCMs ($j_{CO_2}$; Eq. XX), a fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:

@@

where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see Eq. XX). Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress at high light levels, and the emergent outcome of this SU formulation demonstrates a classic photoinhibition response (Fig. S2).

@@

Carbon fixed by photosynthesis ($j_{CP}$; Eq. XX) is then combined with nitrogen shared by the host ($\rho_N$; Eq. XX) and nitrogen recycled from symbiont biomass turnover

@@

Overall, relative changes in the steady state of the system are less than the equivalent relative change in parameter value, indicating that the system maintains internal stability. However, changes in certain parameter values have more significant impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth (Fig. 3), demonstrating the strong nitrogen-limitation that characterizes these symbioses (REF). The parameter $\bar{a}^*$ has a strong impact on S:H biomass ratios (Fig. 3) since this parameter determines the amount of light absorbed by symbionts, with lower values increasing light-limitation. Increasing the maximum growth and turnover rates have the expected effects of increasing and decreasing growth, respectively. Parameters relating to photooxidative stress and bleaching have little impact under these environmental conditions (Fig. 3), but have larger impacts under higher light (e.g., Fig. S6). Sensitivity analyses conducted under different combinations of external light and nutrients are presented in Figs. S3-S7.

>>>>

| Symbol  | Description                            | Units               | Eq. no. |
| ------- | -------------------------------------- | ------------------- | ------- |
| $j_X$     | Prey uptake rate                       | molX CmolH^-1^ d^-1^ | 3 |
| $j_N$     | Nitrogen uptake rate                   | molN CmolH^-1^ d^-1^ | 4 |
| $j_{HG}$  | Host biomass formation rate            | CmolH CmolH^-1^ d^-1^| 5 |
| $r_{NH}$  | Recycled nitrogen from host turnover   | molN CmolH^-1^ d^-1^ | 6 |
| $\rho_N$  | Nitrogen shared with the symbiont      | molN CmolH^-1^ d^-1^ | 7 |
| $j_{eC}$  | Excess carbon used to activate host CCMs| molC CmolH^-1^ d^-1^ | 8 |
| $j_{CO_2}$| CO~2~ input to photosynthesis          | molCO~2~ CmolH^-1^ d^-1^| 9 |
| $j_L$     | Light absorption rate                  | molph CmolS^-1^ d^-1^| 11 |
| $r_{CH}$  | Recycled CO~2~ from host turnover      | molCO~2~ CmolH^-1^ d^-1^| 12 |
| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | molCO~2~ CmolS^-1^ d^-1^| 13 |
| $j_{CP}$  | Photosynthesis rate                    | molC CmolS^-1^ d^-1^ | 14 |
| $j_{eL}$  | Light energy in excess of photochemistry| molph CmolS^-1^ d^-1^| 15 |
| $j_{NPQ}$ | Total capacity of NPQ                  | molph CmolS^-1^ d^-1^ | 16 |
| $c_{ROS}$ | ROS production proportional to baseline | -- | 17 |
| $r_{NS}$  | Recycled nitrogen from symbiont turnover| molN CmolS^-1^ d^-1^| 18 |
| $j_{SG}$  | Symbiont biomass formation rate        | CmolS CmolS^-1^ d^-1^| 19 |
| $\rho_C$  | Fixed carbon shared with host          | molC CmolS^-1^ d^-1^ | 20 |
| $j_{ST}$  | Symbiont turnover rate                 | CmolS CmolS^-1^ d^-1^| 21 |

Table 2. Model parameters. Derivations of each parameter value along with supporting references are provided in the Supplementary Information.

@@

where $\rho_C$ is fixed carbon shared by the symbiont (see Eq. 20), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover:

@@

CO~2~ arrives at the photosynthesis SU from multiple sources: in addition to the CO~2~ actively supplied by the host through its CCMs ($j_{CO_2}$; Eq. 9), a fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:

@@

where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see Eq. 17). Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress at high light levels, and the emergent outcome of this SU formulation demonstrates a classic photoinhibition response (Fig. S2).

@@

Carbon fixed by photosynthesis ($j_{CP}$; Eq. 14) is then combined with nitrogen shared by the host ($\rho_N$; Eq. 7) and nitrogen recycled from symbiont biomass turnover

@@

Overall, relative changes in the steady state of the system are less than the equivalent relative change in parameter value, indicating that the system maintains internal stability. However, changes in certain parameter values have more significant impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth (Fig. 3), demonstrating the strong nitrogen-limitation that characterizes these symbioses. The parameter $\bar{a}^*$ has a strong impact on S:H biomass ratios (Fig. 3) since this parameter determines the amount of light absorbed by symbionts, with lower values increasing light-limitation. Increasing the maximum growth and turnover rates have the expected effects of increasing and decreasing growth, respectively. Parameters relating to photooxidative stress and bleaching have little impact under these environmental conditions (Fig. 3), but have larger impacts under higher light (e.g., Fig. S6). Sensitivity analyses conducted under different combinations of external light and nutrients are presented in Figs. S3-S7.

### 2016-11-02 JTB_ms/JTB_ms.tex
update Eq. nos.

@@

10

@@

@@

Table 2. Model parameters

@@

where \(\rho_C\) is fixed carbon shared by the symbiont (see Eq. XX),
and \(r_{NH}\) is recycled nitrogen liberated by host biomass turnover:

@@

CO\textsubscript{2} arrives at the photosynthesis SU from multiple
sources: in addition to the CO\textsubscript{2} actively supplied by the
host through its CCMs (\(j_{CO_2}\); Eq. XX), a fixed proportion of
metabolic CO\textsubscript{2} generated by host biomass turnover is
passively available to the photosynthesis SU, according to:

@@

where \(j_{CPm}\) is the maximum specific rate of photosynthesis, and
\(c_{ROS}\) is the relative rate of reactive oxygen species production
(see Eq. XX). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
decline in response to photooxidative stress at high light levels, and
the emergent outcome of this SU formulation demonstrates a classic
photoinhibition response (Fig. S2).

@@

Carbon fixed by photosynthesis (\(j_{CP}\); Eq. XX) is then combined
with nitrogen shared by the host (\(\rho_N\); Eq. XX) and nitrogen
recycled from symbiont biomass turnover

@@

Overall, relative changes in the steady state of the system are less
than the equivalent relative change in parameter value, indicating that
the system maintains internal stability. However, changes in certain
parameter values have more significant impacts: increasing \(j_{Nm}\) or
decreasing \(K_N\) both dramatically increase host growth (Fig. 3),
demonstrating the strong nitrogen-limitation that characterizes these
symbioses (REF). The parameter \(\bar{a}^*\) has a strong impact on S:H
biomass ratios (Fig. 3) since this parameter determines the amount of
light absorbed by symbionts, with lower values increasing
light-limitation. Increasing the maximum growth and turnover rates have
the expected effects of increasing and decreasing growth, respectively.
Parameters relating to photooxidative stress and bleaching have little
impact under these environmental conditions (Fig. 3), but have larger
impacts under higher light (e.g., Fig. S6). Sensitivity analyses
conducted under different combinations of external light and nutrients
are presented in Figs. S3-S7.

@@

The prediction of higher growth when light is reduced indicates that
growth is not limited by low light in winter, but is actually reduced by
excess light in summer, consistent with the experimental findings of
Schutter et al. (2011). Seasonal summertime reductions in tissue biomass
have also been well-documented in the field (Fitt et al. 2000), along
with reductions in net photosynthetic capacity (\textbf{???}).
Importantly, while light alone (at these levels) may drive seasonal
dynamics in the ways discussed, temperature fluctuations may attenuate
or even reverse the effect of light as cooler winters depress
metabolism; thus, the relative magnitude of fluctuation in temperature,
light, and other factors may produce wide variability in the direction
and magnitude of seasonal changes in growth and symbiont abundance,
depending on location and microhabitat. Nevertheless, the seasonal
variability predicted here (Fig. 4) is consistent with experimental and
field observations for corals, and demonstrates the power of this model
to predict dynamic behavior that mechanistically integrates multiple
environmental drivers.

@@

>>>>

8

@@

@@

21

@@

Table 2. Model parameters. Derivations of each parameter value along
with supporting references are provided in the Supplementary
Information.

@@

where \(\rho_C\) is fixed carbon shared by the symbiont (see Eq. 20),
and \(r_{NH}\) is recycled nitrogen liberated by host biomass turnover:

@@

CO\textsubscript{2} arrives at the photosynthesis SU from multiple
sources: in addition to the CO\textsubscript{2} actively supplied by the
host through its CCMs (\(j_{CO_2}\); Eq. 9), a fixed proportion of
metabolic CO\textsubscript{2} generated by host biomass turnover is
passively available to the photosynthesis SU, according to:

@@

where \(j_{CPm}\) is the maximum specific rate of photosynthesis, and
\(c_{ROS}\) is the relative rate of reactive oxygen species production
(see Eq. 17). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
decline in response to photooxidative stress at high light levels, and
the emergent outcome of this SU formulation demonstrates a classic
photoinhibition response (Fig. S2).

@@

Carbon fixed by photosynthesis (\(j_{CP}\); Eq. 14) is then combined
with nitrogen shared by the host (\(\rho_N\); Eq. 7) and nitrogen
recycled from symbiont biomass turnover

@@

Overall, relative changes in the steady state of the system are less
than the equivalent relative change in parameter value, indicating that
the system maintains internal stability. However, changes in certain
parameter values have more significant impacts: increasing \(j_{Nm}\) or
decreasing \(K_N\) both dramatically increase host growth (Fig. 3),
demonstrating the strong nitrogen-limitation that characterizes these
symbioses. The parameter \(\bar{a}^*\) has a strong impact on S:H
biomass ratios (Fig. 3) since this parameter determines the amount of
light absorbed by symbionts, with lower values increasing
light-limitation. Increasing the maximum growth and turnover rates have
the expected effects of increasing and decreasing growth, respectively.
Parameters relating to photooxidative stress and bleaching have little
impact under these environmental conditions (Fig. 3), but have larger
impacts under higher light (e.g., Fig. S6). Sensitivity analyses
conducted under different combinations of external light and nutrients
are presented in Figs. S3-S7.

@@

The prediction of higher growth when light is reduced indicates that
growth is not limited by low light in winter, but is actually reduced by
excess light in summer, consistent with the experimental findings of
Schutter et al. (2011). Seasonal summertime reductions in tissue biomass
have also been well-documented in the field (Fitt et al. 2000), along
with reductions in net photosynthetic capacity (Muller-Parker 1987).
Importantly, while light alone (at these levels) may drive seasonal
dynamics in the ways discussed, temperature fluctuations may attenuate
or even reverse the effect of light as cooler winters depress
metabolism; thus, the relative magnitude of fluctuation in temperature,
light, and other factors may produce wide variability in the direction
and magnitude of seasonal changes in growth and symbiont abundance,
depending on location and microhabitat. Nevertheless, the seasonal
variability predicted here (Fig. 4) is consistent with experimental and
field observations for corals, and demonstrates the power of this model
to predict dynamic behavior that mechanistically integrates multiple
environmental drivers.

@@

\hypertarget{ref-MullerParker:1987cr}{}
Muller-Parker, Gisèle. 1987. ``Seasonal variation in light-shade
adaptation of natural populations of the symbiotic sea anemone
Aiptasiapulchella (Carlgren, 1943) in Hawaii.'' \emph{Journal of
Experimental Marine Biology and Ecology} 112 (2): 165--83.
doi:\href{https://doi.org/10.1016/0022-0981(87)90115-8}{10.1016/0022-0981(87)90115-8}.
